---
title: "FinalProj184"
format: pdf
editor: visual
---

### Exploring Government Expenditure on Education Impact on Education

### Liam Geary, Josiah Harrison, Junwen Zeng

With the recent election here in the US a major topic that is frequently brought up is how to improve the education in this country. A major issue with this debate is how drastically opinions can differ between people, with questions of how education has changed for better or worse, what is causing these changes, and even Here we want to focus on the topic that is discussed the most in this debate, which is how increasing funding impacts education.

This is relativity common belief held by many Americans, which is why in this report we will be using data from the World Education Data Set to explore the relationship between


Fork Repository test test


### Literature Review

Install the "readr" Package and get the data from my .csv file.

```{r}
install.packages("readr")
library(readr)

data <- read_csv("/Users/junwen/Desktop/Stat\ 184/world-education-data.csv")
```

Summarize the data to keep the values I need, and also remove the year to keep the mean value in terms of each country.

```{r}
install.packages("dplyr")
library(dplyr)


summarized_data <- data %>%
  group_by(country) %>%
  summarize(
    gov_exp_pct_gdp = mean(gov_exp_pct_gdp, na.rm = TRUE),
    pri_comp_rate_pct = mean(pri_comp_rate_pct, na.rm = TRUE),
    lit_rate_adult_pct = mean(lit_rate_adult_pct, na.rm = TRUE),
    pupil_teacher_primary = mean(pupil_teacher_primary, na.rm = TRUE),
    pupil_teacher_secondary = mean(pupil_teacher_secondary, na.rm = TRUE),
    school_enrol_primary_pct = mean(school_enrol_primary_pct, na.rm = TRUE)
  )

```
The first try of the table, I tried to color-code each country, but due to color-coding all of the countries, each country has a legend label, and the graph is nowhere to be seen, the legend labels even crashed the R program.

```{r}
install.packages("ggplot2")
library(ggplot2)

# 1st
ggplot(summarized_data, aes(x = gov_exp_pct_gdp, y = lit_rate_adult_pct, color = country)) +
  geom_point(size = 3) +
  labs(
    title = "Government Expenditure(In percentage of GDP) vs Literacy Rate by Country",
    x = "Government Expenditure",
    y = "Literacy Rate"
  ) +
  theme_minimal()

```
After removing the legend labels, the color code became meaningless.
```{r}
# 2nd
ggplot(summarized_data, aes(x = gov_exp_pct_gdp, y = lit_rate_adult_pct)) +
  geom_point(size = 3) +
  labs(
    title = "Government Expenditure(In percentage of GDP) vs Literacy Rate by Country",
    x = "Government Expenditure",
    y = "Literacy Rate"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```
The third visualization looks the best, it is quite tidy. This is the visualization of Government Expenditure(In percentage of GDP) vs Literacy Rate by Country.
```{r}
# 3rd
ggplot(summarized_data, aes(x = gov_exp_pct_gdp, y = lit_rate_adult_pct)) +
  geom_point(color = "black", size = 3) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  labs(
    title = "Government Expenditure(In percentage of GDP) vs Literacy Rate by Country",
    x = "Government Expenditure",
    y = "Literacy Rate"
  ) +
  theme_minimal() +
  scale_y_continuous(limits = c(30, 100))

```
This is the visualization of Government Expenditure(In percentage of GDP) vs Primary School Completion Rate Percentage by Country.
```{r}
ggplot(summarized_data, aes(x = gov_exp_pct_gdp, y = pri_comp_rate_pct)) +
  geom_point(color = "black", size = 3) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  labs(
    title = "Government Expenditure(In percentage of GDP) vs Primary School Completion Rate Percentage by Country",
    x = "Government Expenditure",
    y = "Primary School Completion Rate"
  ) +
  theme_minimal() +
  scale_y_continuous(limits = c(30, 100))
```
This is the visualization of Government Expenditure(In percentage of GDP) vs Primary School Pupil Teacher Ratio by Country.
```{r}
ggplot(summarized_data, aes(x = gov_exp_pct_gdp, y = pupil_teacher_primary)) +
  geom_point(color = "black", size = 3) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  labs(
    title = "Government Expenditure(In percentage of GDP) vs Primary School Pupil Teacher Ratio by Country",
    x = "Government Expenditure",
    y = "Primary School Pupil Teacher Ratio"
  ) +
  theme_minimal()
```
This is the visualization of Government Expenditure(In percentage of GDP) vs Secondary School Pupil Teacher Ratio by Country.
```{r}
ggplot(summarized_data, aes(x = gov_exp_pct_gdp, y = pupil_teacher_secondary)) +
  geom_point(color = "black", size = 3) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  labs(
    title = "Government Expenditure(In percentage of GDP) vs Secondary School Pupil Teacher Ratio by Country",
    x = "Government Expenditure",
    y = "Secondary School Pupil Teacher Ratio"
  ) +
  theme_minimal()
```
This is the visualization of Government Expenditure(In percentage of GDP) vs Primary School Enrollment Rate by Country.
```{r}
ggplot(summarized_data, aes(x = gov_exp_pct_gdp, y = school_enrol_primary_pct)) +
  geom_point(color = "black", size = 3) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  labs(
    title = "Government Expenditure(In percentage of GDP) vs Primary School Enrollment Rate by Country",
    x = "Government Expenditure",
    y = "Primary School Enrollment Rate"
  ) +
  theme_minimal()
```
